---
title: "Spatially biased sampling and CKMR"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
```

### Parent-offspring pairs from SLiM

```{r}
parents <- read.csv("parents.txt")
```

### Bravington estimators

Pair $(i, j)$

$K_{ij}$ kinship of the pair

Ages $(y_i, y_j)$

Reproductive output of $j$ at time $i$ $R_i(y_j)$

$\mathbf{z_i} = (t_i, a_i)$ =  (time of sampling, age at sampling)

Age at maturity $\alpha$

$N_{Fy_j}$ number of mature females at time $y_j$ was born

\[
\mathbb{P}(K_{ij} = MO|z_i, z_j) = \frac{\mathbb{E}[R_i(y_j)|y_i, t_i]}{\mathbb{E}[R_+(y_j)]} = \frac{\mathbb{I}[y_i + \alpha \leq y_j < t_i]}{N_{Fy_j}}
\]

## Very simple example

Everyone selfs, so everyone is a mother

$N_{Fy_j} = N$

Since $\alpha = 0$ (everyone is immediately mature), and since time of sampling is the same for everyone

\[
\mathbb{P}(K_{ij} = MO|z_i, z_j) = \frac{\mathbb{I}[y_i < y_j]}{N_{Fy_j}}
\]

In the SLiM simulation not sure if it should be $y_i \leq y_j$ or $y_i < y_j$.

I left it as strictly less than, this also assigns self pairs probability 0 which is good.

```{r}
N <- nrow(parents) # True population size

# Get a random sample of individuals
n <- 200
ind_sample <- parents %>% slice_sample(n = n)

logpslike <- function(N, ind_sample){
  # Log pseudo likelihood
  # All unordered pairs
  pairs <- ind_sample %>% rename(ind_j = individual, parent_of_j = parent1, age_j = age) %>%
    mutate(ind_i = ind_j, age_i = age_j) %>%
    expand(nesting(ind_j, age_j, parent_of_j), nesting(ind_i, age_i)) %>%
    mutate(PO = ind_i == parent_of_j)
  # Probability i is the parent of j for each pair (based on age and N)
  # is 1/N if j is younger than i and 0 otherwise
  pairs <- pairs %>% mutate(P_PO = (age_j < age_i)/N)
  # Log likelihood of observed relationship (not PO or PO)
  pairs <- pairs %>% mutate(loglike = ifelse(PO, log(P_PO), log(1-P_PO)))
  # Pseudo log likelihood is the sum of all the log likelihoods
  return(sum(pairs$loglike))
}

estimate_N <- function(ind_sample){
  # Maximize log pseudo likelihood
  Nhat <- optimize(logpslike, c(0, 2000), tol = 0.0001, maximum = TRUE, ind_sample = ind_sample)$maximum
  return(Nhat)
}

Nhat <- estimate_N(ind_sample)
cat("True N:", N, "Nhat:", Nhat)
```

### Random sampling

```{r}
sample_fun <- function(parents, n){return(slice_sample(parents, n = n))}
n_reps <- 10
n_vals <- seq(10, 300, 50)
n <- rep(n_vals, each = n_reps)
Nhat = map_dbl(n, function(x) estimate_N(sample_fun(parents = parents, n = x)))
random_sampling <- tibble(n = n, Nhat = Nhat)
```

### Spatial sampling

How to sample in a spatially biased way?

Calculate distance to a point, use $1/distance$ as sampling weight.

```{r}
parents <- mutate(parents, sampling_weights = 1/sqrt(x^2 + y^2))

sample_fun <- function(parents, n){return(slice_sample(parents, n = n, weight_by = sampling_weights))}

Nhat = map_dbl(n, function(x) estimate_N(sample_fun(parents = parents, n = x)))
spatially_biased <- tibble(n = n, Nhat = Nhat)
ggplot(spatially_biased, aes(x = n, y = Nhat)) + 
  geom_point() +
  geom_hline(yintercept = N) +
  ggtitle("Spatial")
ggplot(random_sampling, aes(x = n, y = Nhat)) + 
  geom_point() +
  geom_hline(yintercept = N) +
  ggtitle("Random")
```

```{r}
ggplot(parents, aes(x = x, y = y, color = sampling_weights)) +
  geom_point()
```

